---
title: "nanoSIMSiso"
date: "6/26/2021"
output: html_document
---

# N stable isotope probing

Consider a case of a population that is exposed to an isotopically enriched tracer (in our case, this is a $^{15}N $ Nitrogen source). The $^{15}N$ composition of newly enriched biomass will have a distinct isotope composition that reflects tracer addition -- $^{15}F_{B_{new}}$. We can use mass balance between the newly produced and original biomass to determine the overall isotopic composition of biomass at time t:

$$
^{15}F_B (t) = \frac{B_{new}}{B} \cdot \space ^{15}{B_{new}} + \frac{B_{original}}{B} \cdot \space ^{15}F_{B_{original}}
$$
The terms containing $B_{new}$ and $B_{original}$ represent the mass fractions of new and original biomass, respectively.


In the (simplified) case where new biomass mirrors the isotopic composition of the tracer solution (labeled N), we can assume that $^{15}F_{B_{new}} = ^{15}F_{w_{spiked}}$, where $^{15}F_{w_{spiked}}$ is the fractional abundance of $^{15}N$ in the tracer solution.

To calculate apparent turnover time:

$$
F_T = a \cdot F_L \left(1-e^{-\mu\cdot t}\right) + F_{0} \cdot e^{-\mu \cdot t} \\
\downarrow \\
solve \space for \space \mu \\
\downarrow \\
\mu = \frac{1}{t} \cdot \ln\frac{(F_T - a*F_L)}{(F_0 - a*F_L)}
$$
$$
N_t = N_0 \cdot e^{µt} \\
\downarrow \\
\frac{ln(2)}{µ} = t
$$


# Load Libraries

```{r}
library(tidyverse)
library(readxl)
library(ggsci)
library(ggdist)
library(plotly)
library(latex2exp)
library(ggtext)
```

# Import Data and Set Constants

```{r, message=FALSE}
roi_tbl <- read_tsv("roi_tbl.tsv")

a = 0.5 # EDIT: assimilation efficiency
f_l =  1 # EDIT: Labeling strength as 15N frac abundance
t_d = 4 # labeling time in days
nat_abund_15N = .00364 # EDIT: Natural abundance of target isotope 15N frac abundance
f_0 = .00364 # EDIT: Initial Isotopic enrichment 15N frac abundance
```

# FT calculations

Define a function gencalc that calculates growth rate based on assimilation efficiency (a), tracer strength (f_t), initial isotopic enrichment (f_0), label strength (f_l), and incubation time in days (t_d). Returns a growth rate in days.
```{r}
gencalc <- function(a, f_t, f_0, f_l, t_d) {
  case_when(
    t_d == 0 ~ NA_real_,
    TRUE ~ - (1/t_d) * (log((f_t - a*f_l)/(f_0 - a*f_l)))
  )
}
```

```{r}
# Calculate Generation Times
roi_tbl <- roi_tbl %>% mutate(
  mu_d = gencalc(a, Ratio_15N_12Cx14N_12C, f_0, f_l, t_d),
  gen_d = log(2) / mu_d,
)

# Replace negative values with zeros
# If a cell is below natural abundance it will have a "negative" growth rate!
roi_tbl$mu_d[roi_tbl$mu_d < 0] <- 0
roi_tbl$gen_d[roi_tbl$gen_d < 0] <- 0

# Remove NA containing rows
roi_tbl <- roi_tbl %>% drop_na(mu_d)

# Calculate mean turnover and gentimes
roi_tbl <- roi_tbl %>% group_by(N_source, light) %>% 
  mutate(median_mu_d = median(mu_d),
         median_gen_d = median(gen_d))

write_tsv(roi_tbl, "roi_tbl_gentimes.tsv")
```


# Plot

```{r}
p_turnover_nh4_light <- roi_tbl %>% filter(N_source == "NH4", light == "light") %>% 
  ggplot(aes(x = N_source, y = gen_d, color = cell_type)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position=position_jitterdodge(jitter.width = 0.1), size = 0.2) +
  #facet_wrap(vars(light), scales="free_y") +
  ylab(TeX("Cell-Specific Generation Time (days)")) +
  xlab("") +
  scale_color_aaas() +
  coord_cartesian(ylim = c(0, 20)) +
  theme_classic() +
  theme(axis.title.x = element_blank())
p_turnover_nh4_light
```

```{r}
p_nh4_no3 <- roi_tbl %>%
  filter(!is.na(gen_d)) %>% 
  filter(N_source %in% c("no3", "NH4")) %>% 
  ggplot(aes(x = N_source, y = mu_d, color = cell_type)) + 
  #facet_wrap(vars(N_source, light), scale = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position=position_jitterdodge(jitter.width = 0.1), size = 0.2) +
  facet_wrap(vars(light, N_source), scales = "free") +
  ylab(TeX("Cell Turnover $d^{-1}$")) +
  scale_color_aaas() +
  theme_classic()
p_nh4_no3
```

```{r}
p_no_nh4 <- roi_tbl %>%
  filter(!is.na(gen_d)) %>% 
  filter(N_source != "NH4") %>% 
  ggplot(aes(x = N_source, y = mu_d, color = cell_type)) + 
  #facet_wrap(vars(N_source, light), scale = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position=position_jitterdodge(jitter.width = 0.1), size = 0.2) +
  #facet_wrap(vars(light, N_source), scales = "free") +
  ylab(TeX("Cell Turnover $d^{-1}$")) +
  scale_color_aaas() +
  coord_cartesian(ylim = c(0, 0.05)) +
  theme_classic()
p_no_nh4
```


```{r}
roi_tbl %>% filter(cell_type_other == "Picocystis", light == "light") %>% 
  ggplot(aes(x = mu_d, y = N_source)) +
  geom_boxplot(outlier.shape = NA, width = 0.1) +
  geom_jitter(height = 0.1, alpha = 0.5, size = 0.4) +
  ggtitle("Growth under light conditions") +
  theme_classic()
```


```{r}
p_no3 <- roi_tbl %>% filter(
  cell_type_other == "Picocystis", 
  light == "light",
  N_source == "no3") %>% 
  ggplot(aes(x = mu_d, y = N_source)) +
  geom_boxplot(outlier.shape = NA, width = 0.1) +
  geom_jitter(height = 0.1, alpha = 0.5, size = 0.4) +
  ggtitle("Growth under light conditions") +
  theme_classic()
ggplotly(p_no3)
```

```{r}
p_non_nh4 <- roi_tbl %>% filter(
  cell_type_other == "Picocystis", 
  light == "light",
  N_source != "NH4") %>% 
  ggplot(aes(x = mu_d, y = N_source)) +
  geom_boxplot(outlier.shape = NA, width = 0.1) +
  geom_jitter(height = 0.1, alpha = 0.5, size = 0.4) +
  ggtitle("Growth under light conditions") +
  theme_classic()
ggplotly(p_non_nh4)
```

```{r}
roi_tbl %>% filter(
  cell_type_other == "Picocystis", 
  light == "light",
  N_source != "NH4",
  N_source != "no3") %>% 
  ggplot(aes(x = mu_d, y = N_source)) +
  geom_boxplot(outlier.shape = NA, width = 0.1) +
  geom_jitter(height = 0.1, alpha = 0.5, size = 0.4) +
  ggtitle("Growth under light conditions") +
  theme_classic()
```



```{r}
roi_tbl %>% filter(
  cell_type_other == "Picocystis", 
  N_source == "gly") %>% 
  ggplot(aes(x = mu_d, y = N_source)) +
  geom_boxplot(outlier.shape = NA, width = 0.1) +
  geom_jitter(height = 0.1, alpha = 0.5, size = 0.4) +
  facet_wrap(vars(light), ncol=1) +
  ggtitle("Growth with 15N glycine") +
  theme_classic()
```


```{r}
roi_tbl %>% filter(
  cell_type_other == "Picocystis", 
  N_source == "CN") %>% 
  ggplot(aes(x = mu_d, y = N_source)) +
  geom_boxplot(outlier.shape = NA, width = 0.1) +
  geom_jitter(height = 0.1, alpha = 0.5, size = 0.4) +
  facet_wrap(vars(light), ncol=1) +
  ggtitle("Growth with 15N CN") +
  theme_classic()
```


```{r}
roi_tbl %>% filter(
  cell_type_other == "Picocystis", 
  N_source == "TMA") %>% 
  ggplot(aes(x = mu_d, y = N_source)) +
  geom_boxplot(outlier.shape = NA, width = 0.1) +
  geom_jitter(height = 0.1, alpha = 0.5, size = 0.4) +
  facet_wrap(vars(light), ncol=1) +
  ggtitle("Growth with 15N TMA") +
  theme_classic()
```


```{r}
roi_tbl %>% filter(
  cell_type_other == "Picocystis", 
  N_source == "HPG") %>% 
  ggplot(aes(x = mu_d, y = N_source)) +
  geom_boxplot(outlier.shape = NA, width = 0.1) +
  geom_jitter(height = 0.1, alpha = 0.5, size = 0.4) +
  facet_wrap(vars(light), ncol=1) +
  ggtitle("Growth with 15N HPG") +
  theme_classic()
```



```{r}
roi_tbl %>% filter(
  cell_type_other == "Picocystis", 
  N_source == "NH4") %>% 
  ggplot(aes(x = mu_d, y = N_source)) +
  geom_boxplot(outlier.shape = NA, width = 0.1) +
  geom_jitter(height = 0.1, alpha = 0.5, size = 0.4) +
  facet_wrap(vars(light), ncol=1) +
  ggtitle("Growth with 15N NH4") +
  theme_classic()
```

# Remove inactive cells

```{r}
# Take the negative control upper limit (2 sigma)
limit_of_activity <- gencalc(a, .0040436, f_0, f_l, t_d)


roi_tbl_active <- roi_tbl %>% mutate(
  active = case_when(
    mu_d > limit_of_activity ~ "Active",
    TRUE ~ "Inactive"
  )
)

roi_tbl_active_props <- roi_tbl_active %>% 
  group_by(N_source, light) %>% 
  count(active) %>% 
  mutate(prop = n/sum(n)) %>%
  select(-n)

roi_tbl_active_props %>% filter(active == "Active") %>% 
  ggplot(aes(x = N_source, y = prop, fill = light)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_jama() +
  theme_minimal()

roi_tbl_active %>% filter(active == "Active", N_source != "NH4") %>% 
  ggplot(aes(x = N_source, y = mu_d, color = cell_type)) +
  geom_boxplot(outlier.shape = NA) +
  #geom_jitter(size = 0.2) +
  facet_wrap(vars(light), scales = "free") +
  ylab(TeX("Cell Turnover $d^{-1}$")) +
  scale_color_aaas() +
  coord_cartesian(ylim = c(0, 0.05)) +
  theme_classic()
```

```{r}
roi_tbl_active %>% filter(N_source == "NH4", cell_type != "Other") %>% 
  ggplot(aes(x = Ratio_15N_12Cx14N_12C, fill = light)) + 
  stat_slab(alpha = 0.3, height = 0.5) +
  stat_pointinterval(position = position_dodge(width = .4, preserve = "single")) +
  scale_fill_aaas() +
  theme_classic()
```

```{r}
p_NH4_active_raincloud <- roi_tbl_active %>% filter(N_source == "NH4", cell_type != "Other", active == "Active") %>% 
  ggplot(aes(x = Ratio_15N_12Cx14N_12C, fill = light)) + 
  stat_slab(alpha = 0.6, height = 0.5) +
  geom_point(y = -.20, size = 10, shape = 124, alpha = 0.3) +
  stat_pointinterval(y = 0, alpha = 0.5) +
  facet_wrap(vars(light), ncol = 1, strip.position = "right") +
  scale_fill_aaas() +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
ggsave(p_NH4_active_raincloud, filename = "NH4_active_raincloud.png", height = 3, width = 4)
```

```{r}
p_no3_nh4_active_raincloud <- roi_tbl_active %>% 
  filter(N_source == "no3", cell_type != "Other", active == "Active") %>%
  ggplot(aes(x = Ratio_15N_12Cx14N_12C, 
             fill = light, 
             color = light)) +
  stat_slab(alpha = 0.6, height = 0.5) +
  stat_slab(data = roi_tbl_active %>% filter(N_source == "NH4", 
                                             cell_type != "Other", 
                                             active == "Active",
                                             light == "dark"), 
            color = "darkblue", alpha = 0.7, fill = "darkblue") +
  geom_point(data = roi_tbl_active %>% filter(N_source == "NH4", 
                                            cell_type != "Other", 
                                            active == "Active", 
                                            light == "dark"), 
           color = "darkblue", y = -.5, size = 10, shape = 124, alpha = 0.3) +
  stat_pointinterval(data = roi_tbl_active %>% filter(N_source == "NH4", 
                                                      cell_type != "Other", 
                                                      active == "Active", 
                                                      light == "dark"), 
                     color = "black", 
                     alpha = 0.5) +
  geom_point(y = -.5, 
             size = 10, 
             shape = 124, 
             alpha = 0.3) +
  stat_pointinterval(alpha = 0.5, 
                     color = "black") +
  facet_wrap(N_source~light, 
             ncol = 1, 
             strip.position = "right") +
  scale_fill_manual(values = c("orange", "cyan")) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
p_no3_nh4_active_raincloud

#ggsave(p_no3_nh4_active_raincloud, filename = "p_no3_nh4_active_raincloud.png", height = 3, width = 5)
```

```{r}

color_scale = c("#003049", "#f77f00")
p_nh4_no3 <- roi_tbl_active %>% filter(N_source %in% c("NH4", "no3"),
                                       active == "Active") %>% 
  ggplot(aes(x = Ratio_15N_12Cx14N_12C,
             y = N_source,
             fill = light,
             color = light)) +
  stat_halfeye(data = roi_tbl_active %>% filter(N_source == "no3"),
               position = "dodge", 
               color = "black",
               scale = 1) +
  stat_halfeye(data = roi_tbl_active %>% filter(N_source == "NH4"),
               position = "dodge", 
               color = "black",
               scale = 1) +
  geom_point(
    shape = 124,
    size = 5,
    alpha = 0.5,
    position = position_nudge(y = -.3)) +
  facet_wrap(N_source~light, ncol = 1, scale = "free_y") +
  scale_fill_manual(values = color_scale) +
  scale_color_manual(values = color_scale) +
  scale_x_log10() +
  annotation_logticks(sides = "b") +
  theme_classic() +
  theme(
    legend.position = "none",
    panel.grid = element_line(size = 0.2, color = "#ced4da"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    strip.text.y = element_blank(),
    axis.title.x = element_markdown(size = 8.2, margin=margin(t=8)),
    axis.title.y = element_blank(),
    axis.line.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.caption = element_text(size = 7.5, color="#495057"),
    plot.title.position = "plot",
    plot.title = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(size = 8),
    plot.margin = unit(c(.5, .5, .5, .5), "cm")) +
  labs(x = "15N / 14N")
p_nh4_no3

ggsave("nh4LvD.png", p_nh4_no3, width = 4, height = 5, dpi = 300)
```



```{r}
roi_tbl %>% filter(N_source == "no3") %>% 
  group_by(light) %>% 
  summarize(med_mu_d = median(mu_d),
            med_gen_d = median(gen_d))
```

```{r}
p_DON_active_raincloud <- roi_tbl_active %>% 
  filter(N_source %in% c("TMA", "gly", "HPG", "CN"), 
         cell_type != "Other", 
         active == "Active") %>% 
  ggplot(aes(x = Ratio_15N_12Cx14N_12C)) + 
  stat_slab(aes(color = light, fill = light), 
            fill = "gray", 
            alpha = 0.6, 
            height = 0.5) +
  geom_point(aes(color = light), 
             y = -.50, 
             size = 7, 
             shape = 124, 
             alpha = 0.3) +
  stat_pointinterval(y = 0, alpha = 0.5) +
  geom_vline(xintercept = .0040436,
             color = "red",
             alpha = 0.5) +
  facet_wrap(N_source~light, 
             ncol = 1, 
             strip.position = "right") +
  scale_color_aaas() +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
p_DON_active_raincloud
ggsave(p_DON_active_raincloud, filename = "p_DON_active_raincloud.png", height = 5, width = 5)
```

```{r}
median_gen_mu_d <- roi_tbl %>% 
  group_by(N_source, light) %>% 
  summarize(med_gen_d = median(gen_d),
            med_mu_d = median(mu_d),
            mean_gen_d = mean(gen_d)) %>% 
  write_tsv(file = "median_gen_mu_d.tsv")

```

